Compositional equivalences based on Open pNets

### 2020-07-03 2016-Weak-FH-Bisim/WeakBisim.tex
Update WeakBisim.tex
ooops, clean annexe D

\subsubsection*{Running example}
In Section~\ref{sec:weak} we have shown the full satured weak automaton for both the specification and the implementation of the simple protocol. We will show here how we can check if some given relation between these two automata is a weak FH-Bisimulation.

@@

The first piece of code is the textual definition of the SimpleProtSec
pNet, that was drawn in Figure \ref{SimpleProt:Impl}, page \pageref{SimpleProt:Impl}. This code should be intuitive enough to read, with the following
conventions, that brings some user-friendly features that are maps by the editor into pure pNet constructs.

\begin{itemize}
  \item Constants of any type (including Action) must be declared as
    ``const''. They are used either as functions with argument, as
    typically \texttt{in(msg)}, or constants without argument, denoted
    as \texttt{\$tau}. Identifiers without the ``?'' marker are
    variables.
  \item Variables can be declared either: inside a pLTS state (normal
    state variable), as global variable of a pLTS (see e.g. ...), or
    as an input variable in a transition, as \texttt{?msg} in pLTS M1
    here.
  \item the variables in the guards of synchronisation vectors do not need to be explicitely quantified: all variable in a guard that does not appear inside the vector actions will be recognised as bound by a \emph{foral} quantifier inside the guard.
    \item the tools will check that everything is correctly declared, and that variables are used properly, that vectors have coherent length, etc.
\end{itemize}

@@

\begin{lstlisting}[basicstyle=\scriptsize\ttfamily, language=java, frame=single]
SimpleProtImpl2:
import "Data_Alg.algp"
root SimpleProtImpl2

const in,out:Action
const tau,p_send,q_recv,m_recv,m_send,m_error: Action
const s_recv,s_send,s_ack,s_error,r_recv,r_ack,r_send: Action

pLTS M1
  initial a0
  vars ?msg:Data ?c:Int
vars m_msg:Data m_ec:Int

state m0
transition m_recv(msg,c) -> m1 {m_msg:=msg, m_ec:=c}

state m1
transition m_send(m_msg, m_ec) -> m0 
transition synchro($tau) -> m2

state m2
transition $m_error -> m0

pLTS Sender
  initial s0
  vars ?msg:Data
vars  s_msg:Data s_ec:Int

state s0
transition s_recv(msg) -> s1 {s_msg:=msg, s_ec:=0}

state s1
transition s_send(s_msg, s_ec) -> s2

state s2
transition $s_ack -> s0
transition $s_error -> s1 {s_ec:=s_ec+1}

pLTS Receiver
  initial r0
  vars ?msg:Data ?c:Int
vars r_msg:Data r_ec:Int

state r0
transition r_recv(msg,c) -> r1 {r_msg:=msg, r_ec:=c}

state r1
transition r_send(r_msg, r_ec) -> r2

state r2
transition $r_ack -> r0

@@

It is based on the observation that states 202 and 000 are only linked by a "pure $\tau$" transition, and have exactly the same possible behaviours.
More tricky, they do not have the same state variables, as 202 has the variables of state $s2$, but none of these variables are used in transitions from 202.
In this configuration we can guarantee that they are Weak bisimilar, and merge their (incoming and outgoing) transitions in the Figure. We denote this 
equivalence class of states as $\{000,202\}$.

@@

All the following transitions are parametrized by an integer $n\in \mathcal{N}$, meaning they stand for the (infinite) set of weak OTs. 
In some cases, this set is further restricted (see e.g. $WI_{7b}(n)$), in which case we have added an explicit quantifier.
\smallskip

@@

$ WI_4P(n) = \openrule
         {\{\texttt{P}\mapsto \texttt{p-a}\}, [\forall \texttt{m}. \texttt{p-a} \neq \texttt{p-send(m)} ], 
   (\texttt{m\_msg}\gets \texttt{s\_msg}, \texttt{m\_ec}\gets \texttt{s\_ec}+n, \texttt{s\_ec}\gets \texttt{s\_ec}+n)}
         {100 \OTWeakarrow {\texttt{p-a}} 210}
$
and
$ WI_4Q(n) = \openrule
         {\{\texttt{Q}\mapsto \texttt{q-b}\}, [\forall \texttt{m,ec}. \texttt{q-b} \neq \texttt{q-recv(m,ec)} ],
   (\texttt{m\_msg}\gets \texttt{s\_msg}, \texttt{m\_ec}\gets \texttt{s\_ec}+n, \texttt{s\_ec}\gets \texttt{s\_ec}+n)}
         {100 \OTWeakarrow {\texttt{q-b}} 210}
$

@@

We will check here 2 of these conditions.
Preliminary remarks:
  \begin{itemize}
    \item Both pNets trivialy verify the ``non-observability''
      condition: the only vectors having $\tau$ as an action of a
      sub-net are of the form ``$< -, \tau, -> -> \tau$''.
    \item We must take care of variable name conflicts: in our example, the variables of the 2 systems already have different names, but the action parameters occurring in the 
transitions (m, msg, ec) are the same, that is not correct. In the tools, this is managed by the static semantic layer; in the following example, we have renamed all conflicting 
variables using suffix 1 for the Specification, and 2 for the Implementation.
    \end{itemize}

  \medskip
Let us start with the first pair: $SS_1 << WI_1$, where:

@@

  Let check formally the conditions:

@@

\medskip That is trivialy true, choosing pa-2 = pa-1 for each p-a1.

@@

  \bigskip
  All others pairs from this set are just as trivialy prove true.

>>>>

\subsubsection*{Running example}
\label{subsubsection:runnig example}
In Section~\ref{sec:weak} we have shown the full satured weak automaton for both the specification and the implementation of the simple protocol. We will show here how we can check if some given relation between these two automata is a weak FH-Bisimulation.

@@

The first piece of code is the textual definition of the SimpleProtSpec
pNet, that was drawn in Figure \ref{SimpleProt:Impl}, page \pageref{SimpleProt:Impl}. This code should be intuitive enough to read, with the following
conventions, that brings some user-friendly features that are mappeds by the editor into pure pNet constructs.

\begin{itemize}
  \item Constants of any type (including Action) must be declared as
    ``const''. They are used either as functions with argument, as
    typically \texttt{in(msg)}, or constants without argument, denoted
    as \texttt{\$tau}. Identifiers without the ``?'' marker are
    variables.
  \item Variables can be declared as global variables of a pLTS (e.g. \texttt{m\_msg} in \texttt{M1}), or a pNet Node in the case of synchronisation vector variables (e.g. \texttt{p\_a}),
    or as input variables in a pLTS, as \texttt{?msg} in \texttt{M1}.
  \item the variables in the guards of synchronisation vectors do not need to be explicitely quantified: all variables in a guard that do not appear inside the vector actions will be recognised as bound by a \emph{forall} quantifier inside the guard.
    \item the tools will check that everything is correctly declared, and that variables are used properly and do not conflict between different objetcs, that vectors have coherent length, etc.
\end{itemize}

@@

\begin{lstlisting}[basicstyle=\scriptsize\ttfamily, language=java, frame=single]
SimpleProtImpl2:
  import "Data_Alg.algp"
  root SimpleProtImpl2
const in,out:Action
const tau,p_send,q_recv,m_recv,m_send,m_error: Action
const s_recv,s_send,s_ack,s_error,r_recv,r_ack,r_send: Action

pLTS M1
  initial a0
  vars ?msg:Data ?c:Int
  vars m_msg:Data m_ec:Int
state m0
  transition m_recv(msg,c) -> m1 {m_msg:=msg, m_ec:=c}
state m1
  transition m_send(m_msg, m_ec) -> m0 
  transition synchro($tau) -> m2
state m2
  transition $m_error -> m0

pLTS Sender
  initial s0
  vars ?msg:Data
  vars  s_msg:Data s_ec:Int
state s0
  transition s_recv(msg) -> s1 {s_msg:=msg, s_ec:=0}
state s1
  transition s_send(s_msg, s_ec) -> s2 
state s2
  transition $s_ack -> s0
  transition $s_error -> s1 {s_ec:=s_ec+1}

pLTS Receiver
  initial r0
  vars ?msg:Data ?c:Int
  vars r_msg:Data r_ec:Int
state r0
  transition r_recv(msg,c) -> r1 {r_msg:=msg, r_ec:=c}
state r1
  transition r_send(r_msg, r_ec) -> r2
state r2
  transition $r_ack -> r0

@@

This drawing is based on the observation that states 202 and 000 are only linked by a "pure $\tau$" transition, and have exactly the same possible behaviours.
More tricky, they do not have the same state variables, as 202 has the variables of state $s2$, but none of these variables are used in transitions from 202.
In this configuration we can guarantee that they are Weak bisimilar, and we have merged their (incoming and outgoing) transitions in the Figure. We denote this 
equivalence class of states as $\{000,202\}$.

@@

All the following transitions are parametrized by an integer $n\in \mathcal{N}$, meaning they stand for the corresponding (infinite) set of weak OTs. 
In some cases, this set is further restricted (see e.g. $WI_{7b}(n)$), in which cases we have added an explicit quantifier.
\smallskip

@@

$ WI_4P(n) = \openrule
         {\{\texttt{P}\mapsto \texttt{p-a}\}, [\forall \texttt{m}. \texttt{p-a} \neq \texttt{p-send(m)} ], \\
   (\texttt{m\_msg}\gets \texttt{s\_msg}, \texttt{m\_ec}\gets \texttt{s\_ec}+n, \texttt{s\_ec}\gets \texttt{s\_ec}+n)}
         {100 \OTWeakarrow {\texttt{p-a}} 210}
$

and
$ WI_4Q(n) = \openrule
         {\{\texttt{Q}\mapsto \texttt{q-b}\}, [\forall \texttt{m,ec}. \texttt{q-b} \neq \texttt{q-recv(m,ec)} ],\\
   (\texttt{m\_msg}\gets \texttt{s\_msg}, \texttt{m\_ec}\gets \texttt{s\_ec}+n, \texttt{s\_ec}\gets \texttt{s\_ec}+n)}
         {100 \OTWeakarrow {\texttt{q-b}} 210}
$

@@

Preliminary remarks:
  \begin{itemize}
    \item Both pNets trivialy verify the ``non-observability''
      condition: the only vectors having $\tau$ as an action of a
      sub-net are of the form ``$< -, \tau, -> -> \tau$''.
    \item We must take care of variable name conflicts: in our example, the variables of the 2 systems already have different names, but the action parameters occurring in the 
transitions (m, msg, ec) are the same, that is not correct. In the tools, this is managed by the static semantic layer; in the following example, we have renamed all conflicting 
variables using suffix 1 for the Specification, and 2 for the Implementation.
    \end{itemize}

  In our running example in page \pageref{subsubsection:runnig example}, we have shown the proof for one of the
transitions of ($\text{b0},  202,  True$), namely that $SS_3$ is covered by $WI_3(0)$.
We give here another example with $SS_1 << WI_1$, from the first triple ($\text{b0},  000,  True$). It includes less trivial predicates in the OTs:

@@

@@

\medskip That is trivialy true, choosing \texttt{p-a1=p-a2} for each given \texttt{p-a1}.

@@

  \bigskip
  All others pairs from this set are just as easily proved true.
